---
title: "genotype"
execute:
  echo: false
  warning: false
  message: false
format: html
---

```{r}
source("functions.R")
```

```{r}
library(tidyverse)
library(reshape2)
library(ggplot2)
library(ggtext)
library(vdjbasevis)
```

```{r}
num_subjects = length(unique(data_$sample))

```


```{r,echo=FALSE}

imgt_genes = 'IGHV'

genes <- unique(bag_data[bag_data$in_baseline_reference=="TRUE","gene"])
ASCs_on_ref= genes[grepl(imgt_genes, genes)]
lenght_ASCs_on_ref=length(ASCs_on_ref)

genes_on_genotype <- unique(bag_data[bag_data$sample_count > 0,"gene"])
ASCs_on_genotype = genes_on_genotype[grepl(imgt_genes, genes_on_genotype)]
novel_ASCs_on_genotype = setdiff(ASCs_on_ref,ASCs_on_genotype)
length_novel_ASCs_on_genotype = length(novel_ASCs_on_genotype)
lenght_ASCs_on_genotype = length(setdiff(ASCs_on_genotype,novel_ASCs_on_genotype) )

lenght_ASCs_not_represented = lenght_ASCs_on_ref - lenght_ASCs_on_genotype

allele <- unique(bag_data[bag_data$in_baseline_reference=="TRUE","allele"])
alleles_on_ref=allele[grepl(imgt_genes, allele)]
lenght_alleles_on_ref=length(alleles_on_ref)

allele_on_genotype <- unique(bag_data[bag_data$sample_count > 0,"allele"])
alleles_on_genotype = allele_on_genotype[grepl(imgt_genes,allele_on_genotype)]
novel_alleles_on_genotype = setdiff(c(alleles_on_genotype),alleles_on_ref)
length_novel_alleles_on_genotype = length(novel_alleles_on_genotype)
lenght_alleles_on_genotype = length(setdiff(c(alleles_on_genotype),novel_alleles_on_genotype) )

lenght_alleles_not_represented = lenght_alleles_on_ref - lenght_alleles_on_genotype
```


For the genotype inference we used the following conservative criteria:

- Zero mutations within the V region.
- Single assignment for thr V.
- Started from the first position of the V gene.


The Piglet allele-based method was used for genotype inference based on AIRR-seq data. It compares allele usage against a population-derived threshold to determine allele inclusion in the genotype. A Z score is calculated for each allele, measuring its significance in the repertoire. The formula incorporates allele count, frequency threshold, and repertoire depth, with higher Z scores indicating a stronger likelihood of allele presence in the genotype.

In a detailed analysis, it was found that of the `r lenght_ASCs_on_ref` existing IGHV clusters, `r lenght_ASCs_not_represented` are not represented in the genotypes of the `r num_subjects` subjects analyzed. `r lenght_alleles_not_represented` out of `r lenght_alleles_on_ref` existing IGHV alleles were not found in the genotypes of the subjects.

There are `r length_novel_alleles_on_genotype` IGHV novel allele that were found in the genotypes of these `r num_subjects` subjects.



```{r}
allele_palette_new <-vdjbasevis:::ALLELE_PALETTE
names(allele_palette_new) <- c("01" ,"02", "03", "04" ,"05", "06", "07","08" ,"09" ,"10", "11", "12", "13", "14", "15",
                               "16" ,"17", "18", "19", "20" ,"21" ,"22" ,"23" ,"24", "25" ,"26", "28", "29", "31", "32","NA")
```


```{r}
data_ = read.csv("actual_df.csv")
data_ <- data_ %>%
      group_by(gene, sample) %>%
        mutate(n = n(), idx = row_number()) %>%
        ungroup()


library(stringr)

data_$genotyped_allele <- sub("^[^*]*\\*", "", data_$allele)
```


```{r}
df_expanded <- data_ %>%
  mutate(subject_idx = as.numeric(factor(sample)),
         gene_idx = as.numeric(factor(gene)))

df_expanded$genotyped_allele <-as.factor(df_expanded$genotyped_allele)

df_expanded<- df_expanded %>%
  mutate(modified_allele = ifelse(grepl("_", genotyped_allele),
                                  sub("_.*", "", genotyped_allele),
                                  as.character(genotyped_allele)))


unique_alleles_with_underscore <- df_expanded %>%
  arrange(genotyped_allele)%>%
  mutate(genotyped_allele = as.factor(genotyped_allele))%>%
  filter(grepl("_", genotyped_allele)) %>%
  distinct(genotyped_allele) %>%
  mutate(underscore_number = seq_along(genotyped_allele))

df_expanded <- df_expanded %>%
  left_join(unique_alleles_with_underscore, by = "genotyped_allele") %>%
  mutate(underscore_number = ifelse(is.na(underscore_number), NA_integer_, underscore_number))

df_expanded <- df_expanded %>%
  mutate(has_underscore = grepl("_", genotyped_allele))

legend_data <- df_expanded %>%
  distinct(genotyped_allele, underscore_number) %>%
  mutate(legend_label = ifelse(is.na(underscore_number),
                               as.character(genotyped_allele),
                               as.character(paste( "^" ,underscore_number, "^" ,genotyped_allele, sep = "")))) %>%
  arrange(genotyped_allele)


unique_alleles <- unique(df_expanded$genotyped_allele)
unique_modified_alleles <- unique(df_expanded$modified_allele)

# Initialize allele_palette_modified with colors from allele_palette_new
allele_palette_modified <- allele_palette_new

# Loop through the unique_modified_alleles to assign colors
for (allele in unique_alleles) {
  if (!(allele %in% names(allele_palette_new))) {
       allele_palette_modified[allele] <-  allele_palette_new[sub("_.*", "", allele)]
  } else {
    
    allele_palette_modified[allele] <- allele_palette_new[allele]
  }
}

```

```{r}
p<-ggplot(df_expanded, aes(xmin = subject_idx - 0.5 + (idx-1)/n, xmax = subject_idx - 0.5 + idx/n,
                        ymin = gene_idx - 0.5, ymax = gene_idx + 0.5)) +
  geom_rect(aes(fill = genotyped_allele, text = paste("</br>genotyped_allele : ",genotyped_allele,
                                                      "</br>Subject: ", sample))) +
  geom_text(data = subset(df_expanded, has_underscore), 
            aes(x = subject_idx - 0.5 + (idx-0.5)/n, y = gene_idx, label =   paste( "^" ,underscore_number, sep = "")), 
            vjust = 0.5, hjust = 0.5, 
            size = 2)  +
  scale_fill_manual(values = allele_palette_modified, name = " ",
                    breaks = legend_data$genotyped_allele,
                    labels = legend_data$legend_label) +
  scale_x_continuous(breaks = df_expanded$subject_idx, labels = df_expanded$sample,
                     expand = c(0, 0)) +  # no expansion
  scale_y_continuous(breaks = df_expanded$gene_idx, labels = df_expanded$gene,
                     expand = c(0, 0)) +
  theme_minimal() +
  geom_vline(xintercept = unique(df_expanded$subject_idx) - 0.5, color = "black", size = 0.2) +
  geom_hline(yintercept = unique(df_expanded$gene_idx) - 0.5, color = "black", size = 0.2) +
  labs(fill = " ", x = " ", y = " ") +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5, size = 10),
        axis.text.y = element_text(angle = 0, hjust = 0.5, size = 10),
        legend.position = "none",
        legend.key.size = unit(0.3, "cm"))

```

